---
execute: 
  echo: false
---

```{r}
#| message: false
library(tidyverse)
library(plotly)
```

```{r}
# Read Import/Export plot of digital services
plot <- readRDS("data/main_plot.rds")

# Show plot with plotly
ggplotly(plot)
```

Since digital devices are everywhere, most businesses need online services to stay connected. Using data from the UN Conference on Trade and Development (UNCTAD) and the World Bank, covering country-level statistics and indicators on the digital economy from 2010 to 2023, particularly in developing countries. I used R to look at the insights into how the digital economy and technology are contributing to global trade (Especially, imports and exports of digital services). The digital economy has seen robust growth over the past decade, with digitally-deliverable services trade nearly doubling between 2010 and 2023. It is a consistent upward trend that reflects the increasing integration of digital services into the global economy, now accounting for over 7.5% of global GDP.