---
execute: 
  echo: false
---

```{r}
#| message: false
library(tidyverse)
library(plotly)
library(RColorBrewer)
```

```{r}
# Read dataset
df <- read_csv("data/UNCTAD_DE.csv", show_col_types = FALSE)
```

## Export distribution by country (2023)

```{r}
# Transforming exports rows
exports_pie_data <- df |>
  filter(
    INDICATOR == "UNCTAD_DE_DIG_SERVTRADE_ANN_EXP",
    TIME_PERIOD == 2023,
    REF_AREA != "WLD", # Get rows for individual countries
    !is.na(OBS_VALUE)
  ) |>
  group_by(REF_AREA_LABEL) |>
  summarise(
    total_export = sum(OBS_VALUE * 10^UNIT_MULT / 1e9, na.rm = TRUE)
  ) |>
  arrange(desc(total_export)) |>
  mutate(
    rank = row_number(),
    group = ifelse(rank <= 10, REF_AREA_LABEL, "Others")
  ) |>
  group_by(group) |>
  summarise(total_export = sum(total_export)) |>
  mutate(
    percentage = total_export / sum(total_export) * 100,
    value_billion = total_export, # assumes OBS_VALUE is already in billions
    label = paste0(
      group,
      ": $",
      round(value_billion, 1),
      "B (",
      round(percentage, 1),
      "%)"
    ),
    pull_out = if_else(group == "Others", 0.1, 0) # pull "Others" slice
  )
```

```{r}
# Interactive pie chart with better layout
plot_ly(
  data = exports_pie_data,
  labels = ~group,
  values = ~value_billion,
  type = 'pie',
  text = ~label,
  textinfo = 'text',
  hoverinfo = 'text',
  pull = ~pull_out,
  marker = list(
    colors = brewer.pal(n = min(11, nrow(exports_pie_data)), name = "Set3")
  ),
  insidetextorientation = 'radial'
) |>
  layout(
    title = list(text = "Top 10 Exporting Countries (2023)", x = 0.5),
    showlegend = FALSE,
    margin = list(t = 50, b = 50, l = 50, r = 50)
  )
```

The United States leads the list with $680.5 billion in exports, representing 15.7% of total exports among the top countries. The United Kingdom follows with $449.2 billion (10.3%), and Ireland stands at $340.1 billion (7.8%). India, Germany, and China also contribute significantly, ranging from 5% to 6.6% each. Notably, while these top 10 countries collectively hold a major share, the "Others" category accounts for the largest portion at $1,409 billion or 32.4%, indicating that a substantial portion of global exports comes from countries outside the top 10. This highlights both the dominance of key economies and the diverse participation of smaller nations in global trade.

## Import distribution by country (2023)

```{r}
# Transforming imports rows
imports_pie_data <- df |>
  filter(
    INDICATOR == "UNCTAD_DE_DIG_SERVTRADE_ANN_IMP",
    TIME_PERIOD == 2023,
    REF_AREA != "WLD", # Get rows for individual countries
    !is.na(OBS_VALUE),
    !is.na(UNIT_MULT)
  ) |>
  mutate(value_billion = OBS_VALUE * 10^UNIT_MULT / 1e9) |> # Convert to Billions USD
  group_by(REF_AREA_LABEL) |>
  summarise(total_import = sum(value_billion, na.rm = TRUE)) |>
  arrange(desc(total_import)) |>
  mutate(
    rank = row_number(),
    group = ifelse(rank <= 10, REF_AREA_LABEL, "Others")
  ) |>
  group_by(group) |>
  summarise(total_import = sum(total_import)) |>
  mutate(
    percentage = total_import / sum(total_import) * 100,
    label = paste0(
      group,
      ": $",
      round(total_import, 1),
      "B (",
      round(percentage, 1),
      "%)"
    ),
    pull_out = if_else(group == "Others", 0.1, 0)
  )
```

```{r}
# Create Plotly pie chart
plot_ly(
  data = imports_pie_data,
  labels = ~group,
  values = ~total_import,
  type = 'pie',
  text = ~label,
  textinfo = 'text',
  hoverinfo = 'text',
  pull = ~pull_out,
  marker = list(
    colors = brewer.pal(n = min(11, nrow(imports_pie_data)), name = "Set3")
  ),
  insidetextorientation = 'radial'
) |>
  layout(
    title = list(text = "Top 10 Importing Countries (2023)", x = 0.5),
    showlegend = FALSE,
    margin = list(t = 50, b = 50, l = 50, r = 50)
  )
```

The United States ranks first with imports totaling $409.7 billion, accounting for 11.1% of the total. Ireland follows closely at $361.8 billion (9.8%), with Germany at $256.6 billion (7%) and the United Kingdom at $218.2 billion (5.9%). Other notable importers include the Netherlands, France, China, Singapore, Japan, and Switzerland, each contributing between 3.4% and 5.6%. Despite these significant figures, the "Others" category dominates, representing $1,455.9 billion or 39.4% of total imports. This underscores the broad distribution of global import activity, with numerous countries outside the top 10 playing a vital role in international trade.
